An Embedding Framework for Consistent Polyhedral Surrogates

### 2019-05-23 neurips-19/poly-embed.tex
Clarifying top k
Initial changes to top k post discussion

\subsection{Inconsistency of top-$k$ losses}
~\citet{lapin2016loss, lapin2018analysis}, provide two surrogate losses for measuring top-$k$ error, one of which is shown to be convex, but their consistency is left as an open question.

The top $k$ classification problem takes as input a set $r \in \R := \{r \subseteq [n] : |r| = k\}$, and observes outcome $y \in [n]$.
We define the top-$k$ loss $\ell_{top-k}:\R \to \reals$ by
\begin{align}\label{eq:top-k}
\ell_{top-k}(r,y) = \begin{cases}
0 & y \in r\\
1 & \text{otherwise}
\end{cases}
\end{align}

Yang and Koyejo~\cite{yang2018consistency} create a new surrogate loss that is consistent for the top-$k$ problem, but it should be noted that their surrogate is not convex.
They also show that Lapin's surrogates are not consistent, but in a different manner than we do.
Following the notation of Yang, we call Lapin's~\cite{lapin2018analysis} convex surrogate $L_2$, and focus on this.
Here, we assume $r \subseteq [n]$ and $k \leq n$ is a constant.

Our embedding framework can help us find distributions where the surrogate loss is not minimized by predicting the minimizer of $\ell_{top-k}$.
First, we want to show that for all $p \in \simplex$, there (1.) is some $u \in \{0,1\}^n$ that minimizes $\inprod{p}{L_2(\cdot)}$.
Then, we show (2.) there is a discrete loss $\ell_2 : \{0,1\}^n \to \reals^Y_+$ that is embedded by $L_2$.  %, where the embedded reports are the corners of the $n$-dimensional hypercube.
For $k=1$, we know that the $\ell_{top-k}$ reduces to 0-1 loss, so finally, we (3.) characterize the distributions where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

\raft{Take another look at Theorem 2.4 of the top-k paper.  Losses can certainly be consistent without embedding $\ell$ (even if polyhedral: you can link multiple embedded points to a single discrete report, e.g.\ a convex surrogate for ranking loss can link to the mode), so the ``iff'' is not right.  I think it's worth spending some time thinking about what exactly we need to show to prove inconsistency.  In particular, see if this reasoning checks out:  Suppose we find a distribution $p$ for which an optimal surrogate report is $r$ but $\psi(r)$ is not optimal for $\ell$.  Since we could have a conditional distribution of $p$ for every $x$, so the optimal hypothesis would just be the constant function $h(x) = r$, which achieves the optimal $L$-loss, and when linked, $\psi \circ h(x) = \psi(r)$, which does not achieve the optimal $\ell$-loss.}

We write the surrogate $L_2(u,y) := \left( \frac{1}{k} \sum_{i=1}^k (u + \indopp(y))_{[i]} - u_y \right)_+ $, where $\indopp(y) = \ones - e_y$, and $(x)_{[i]}$ is the $i^{th}$ largest component of $x$.
$L_2$ is a polyhedral loss, since it can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \indopp(y)$ with highest weight.
By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L_2}$ is also polyhedral.

When embedding, we consider the discrete loss
\begin{align*}
\ell_2(r,y) &= \Ind[y \not\in r] + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)~,~
\end{align*}
and claim that $L_2$ embeds $\ell_2$.

Finally, taking $k = 1$, we see that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$, while the expected loss for $\ell_{top-k}$ is strictly less when predicting $\argmax_y p_y$.

\begin{enumerate}
	\item First, note that for all $p \in \simplex$, there is some $u \in \argmin_{u' \in \reals^n} \inprod{p}{L(u')} \cap [0,1]^n$, since $e_y$ is a minimizer of $\inprod{\delta_y}{L(\cdot)}$ for all $y \in \Y$, so there is a minimizer in $\conv(\{\delta_y\}_{y \in \Y}) = [0,1]^n$.

	Now we proceed in two cases: considering distributions $p$ where $\exists y \in \Y$ so that $p_y \geq 1/2$ and then considering $p$ such that $\max_{y} p_y \leq 1/2$.

	For the first case, we claim that $e_y \in \argmin_{u \in [0,1]^n} \inprod{p}{L_2(u)}$.
	Consider
	\begin{align*}
		L_2(e_y, p) &= \sum_{y' \neq y} p_{y'} L_2(e_y, y') + L_2(e_y, y)\\
		&= \frac{k+1}{k} (1 - p_y) %+ 0 \\
%		&\leq \frac{k+1}{2k} \\
%		&\leq 1
	\end{align*}
	\jessie{TODO-- Not sure how.  Two thoughts: can we leverage Jensen's inequality since we have convexity?  Or can we show these are supporting hyperplanes of the risk?}

	\item We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{0,1\}^n$.
	Moreover, by part (1.) and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \{0,1\}^n} \inprod{p}{L_2(u)} = \min_{r \subseteq}\ell_2(r)$, and so we conclude that $L_2$ embeds $\ell_2$.

	\item First, we see that $\ell(\vec{0}, \bar{p}) = 1$.
	Taking any basis vector $e_y$, the expected loss $\ell_2(e_y, p) = 2 (1-p_y) \geq 1$ by assumption on $p$. 
	For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r,y) \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
	Therefore, we have that $0 \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.
\end{enumerate}

Thus, we can use our embedding framework to distinguish distributions where a surrogate is not consistent with respect to a given discrete loss.

@@

>>>>

\subsection{Inconsistency of top-$k$ losses}

The top-$k$ classification problem takes as input a set $r \in \R := \{r \subseteq [n] : |r| = k\}$, and observes outcome $y \in [n]$.
We define the top-$k$ loss $\ell_{top-k}:\R \to \reals$ by
\begin{align}\label{eq:top-k}
\ell_{top-k}(r,y) = \begin{cases}
0 & y \in r\\
1 & \text{otherwise}
\end{cases}
\end{align}

\citet{lapin2016loss, lapin2018analysis}, provide two surrogate losses for measuring top-$k$ error, one of which is shown to be convex, but their consistency is left as an open question.
\citet{yang2018consistency} show the inconsistency of their losses, and create a new surrogate loss that is consistent for the top-$k$ problem, but it should be noted that Yang and Koyejo's surrogate is not convex.
They also show that Lapin's surrogates are not consistent, but without too much intuition as to why.
Following the notation of Yang, we call Lapin's~\cite{lapin2018analysis} convex surrogate $L_2$, and focus on this.
Here, we assume $r \subseteq [n]$ and $k \leq n$ is a constant.

We show that $L_2$ embeds some other discrete loss $\ell_2:\{0,1\}^n \to \reals^\Y_+$, which has a different Bayes Risk than $\ell_{0-1}$.

In order to do this, we want to show that for all $p \in \simplex$, $\inf_{u' \in \reals^n} \inprod{p}{L_2(u')} = \min_{u' \in \{0,1\}^n} \inprod{p}{L_2(u')}$. (Lemma~\ref{lem:top-k-optimal-corners}.)
Then, we show there is a discrete loss $\ell_2 : \{0,1\}^n \to \reals^\Y_+$ that is embedded by $L_2$.  (Lemma~\ref{lem:top-k-surrogate-embeds}.)

We conclude with a concrete example; $k=1$.
One can see that the $\ell_{top-1}$ reduces to 0-1 loss, and we characterize the distributions where $\risk{L_2}(p) = \risk{\ell_2}(p) \neq \risk{\ell_{top-k}}(p)$.

%\raft{Take another look at Theorem 2.4 of the top-k paper.  Losses can certainly be consistent without embedding $\ell$ (even if polyhedral: you can link multiple embedded points to a single discrete report, e.g.\ a convex surrogate for ranking loss can link to the mode), so the ``iff'' is not right.  I think it's worth spending some time thinking about what exactly we need to show to prove inconsistency.  In particular, see if this reasoning checks out:  Suppose we find a distribution $p$ for which an optimal surrogate report is $r$ but $\psi(r)$ is not optimal for $\ell$.  Since we could have a conditional distribution of $p$ for every $x$, so the optimal hypothesis would just be the constant function $h(x) = r$, which achieves the optimal $L$-loss, and when linked, $\psi \circ h(x) = \psi(r)$, which does not achieve the optimal $\ell$-loss.}

We write the surrogate 
\begin{align*}
L_2(u,y) := \left( \frac{1}{k} \sum_{i=1}^k (u + \ones - e_y)_{[i]} - u_y \right)_+
\end{align*}
where $x_{[i]}$ is the $i^{th}$ largest component of $x$.
By Lemma~\ref{lem:top-k-optimal-corners}, we can see that for all $p \in \simplex$, there is some $u \in \{0,1\}^n$ so that $u \in \argmin_{u' \in \reals^d} \inprod{p}{L_2(u)}$.
With this, we show that $L_2$ actually embeds a discrete loss defined on the corners of the hypercube.

When embedding, we consider the discrete loss $\ell_2: \{0,1\}^n \to \reals^Y_+$
\begin{align*}
\ell_2(r,y) &= \Ind[y \not\in r] + \min\left(\frac{|r \setminus \{y\}|}{k}, 1\right)~,~
\end{align*}
and we show that $L_2$ embeds $\ell_2$ in Lemma~\ref{lem:top-k-surrogate-embeds}.
Moreover, by Proposition~\ref{prop:embed-bayes-risks}, we know $\risk{L_2} = \risk{\ell_2}$.

Finally, taking $k = 1$, we see that $\vec{0} \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$, while the expected loss for $\ell_{top-k}$ is strictly less when predicting $\argmax_y p_y$.
First, we see that $\ell(\vec{0}, \bar{p}) = 1$.
Taking any basis vector $e_y$, the expected loss $\ell_2(e_y, p) = 2 (1-p_y) \geq 1$ by assumption on $p$. 
For any $r$ such that $|r| \geq 2$ \jessiet{In general, $|r| \geq k+1$}, we know that  $\ell_2(r,y) \geq 1$ for all $y \in \Y$, so the expected loss much be greater than or equal to $1$.
Therefore, we have that $0 \in \argmin_{r' \subseteq [n]} \inprod{p}{\ell_2(r')}$ for any $p$ such that $\max_y p_y \leq 1/2$.

So while $L_2$ given by~\citet{lapin2018analysis} is not consistent with respect to top-$k$ loss, we do show a discrete loss $\ell_2$ so that $L_2$ is consistent with respect to $\ell_2$.

@@

\section{Top-$k$}
Throughout this section, consider the surrogate and discrete loss\jessie{fill in}

Observe that $L_2$ is a polyhedral loss, since it can be written as the pointwise max of $\binom{n}{k} +1$ terms, where the $\binom{n}{k}$ terms are selecting the $k$ elements of $u + \indopp(y)$ with highest weight.
By Lemma~\ref{lem:poly-loss-poly-risk}, we then know that $\risk{L_2}$ is also polyhedral.

\begin{lemma}\label{lem:top-k-optimal-corners}
	For all $p \in \simplex$, there is a $u \in \{0,1\}^n$ such that $u \in \argmin_{u'} \inprod{p}{L_2(u)}$.
\end{lemma}
\begin{proof}

\end{proof}

\begin{lemma}\label{lem:top-k-surrogate-embeds}
$L_2:\reals^n \to \reals^\Y_+$ embeds $\ell_2:2^{[n]} \to \reals^\Y_+$.
\end{lemma}
\begin{proof}
  We can verify $\ell_2(r) = L_2(\varphi(r))$ for all $r \subseteq [n]$ and $\varphi(r)$ being its conversion from a set to binary vector, so $\varphi(\R) = \{0,1\}^n$.
  Moreover, by part (1.) and the above equality, respectively, we know that $\inf_{u \in \reals^n} \inprod{p}{L_2(u)} = \min_{u \in \{0,1\}^n} \inprod{p}{L_2(u)} = \min_{r \subseteq}\ell_2(r)$, and so we conclude that $L_2$ embeds $\ell_2$.
\end{proof}
